### 2024-08-29 draft_1.qmd
continued

---
title: "Comparison of Confidence Interval in the OLS and Ridge Linear Regression Model: A Comparative study via Simulation"
format: #docx
    tandf-pdf:
      keep-tex: true  
    tandf-html: default
author:
  - name: Sultana Mubarika Rahman Chowdhury
    affiliations:
      - ref: biostat
    orcid: 
    email: schow034@fiu.edu
  - name: B.M. Golam Kibria
    affiliations:
      - ref: stat
  - name: Zoran Bursac
    affiliations:
      - ref: biostat
affiliations:
  - id: biostat
    name: Florida International University
    department: Biosatistics Department
    address: 11200 SW 8th St
    city: Miami, Fl
    country: USA
    postal-code: 33199
  - id: stat
    name: Florida International University
    department: Department of Mathematics and Statistics
    address: 11200 SW 8th St 
    city: Miami, Fl
    country: USA
    postal-code: 33199
abstract: |
  Write down the abstract here
keywords: 
  - Multiple Linear Regression
  - Ridge Regression
  - Multicolinearity
bibliography: bibliography.bib  
---

# Introduction

Multiple linear regression maps the relationship between two or more predictors and dependent variable to a linear equation. The aim is to predict the response variable using the independent variables. If X is a n×p full rank matrix of predictors and Y is a n × 1 vector of response variables the multiple linear regression can be explained as,

$$Y = X \times \beta + \epsilon, $$ where, $\beta$ is an p × 1 unknown regression paramaters and $\epsilon$ is the n × 1 vector of error with mean zero and equal variance.Ordinary Least Square (OLS) method is commonly used to estimate the unknown regression parameter. The OLS estimate in case of linear regression model is defined as follows,

$$\hat{\beta} = (X^T X)^{-1} X^T y.$$

One of the key assumption of the widely used multiple linear regression model is that the predictors need to be independent of each other. Violating the assumption of independence results in an issue known as multicollinearity. Multicollinearity, originally identified by @frisch1934statistical, is the state in which two or more independent variables show a strong correlation with one another. It causes standard error of the OLS estimator to increase resulting in wide confidence intervals and less reliable results. Identifying and interpreting significant predictors also becomes difficult [@saleh2019theory]. Increasing the sample size, eliminating the highly correlated variables, Principal component analysis are some of the ways to avoid multicollinearity but with the risk of losing either valuable information or interpretability.

As the field of study progressed, researchers developed a number of methods to perform better than OLS in presence of multicollinearity in data by introducing bias to the estimator to gain smaller variance. Ridge regression [@hoerl1970ridge], Lasso [@tibshirani1996regression], Stein estimator [@stein1956inadmissibility], Modified ridge regression, Liu estimator [@liu1993new], 
Kibria-Lukman estimator [@kibria2020new] are some them.

Ridge regression is one of the methods proposed by @hoerl1970ridge that mitigates the effects of multicollinearity by introducing an amount of bias in to the model.

para about differnt techniques,

para about ridge

para about CI

para about Aim

Note:

# Statistical Methodology

## Ridge Regression Estimators

## Confidence Interval

## CI for Ridge regression

# Simulation Study

# Application

# Discussion and Conclusion

# Bibliography